The 2018 heatwave and its implications on ozone induced damage on vegetation in a subarctic climate

### 2020-04-16 VegOzone.tex
Some rearangement of figures and added a new figure.

The \chem{\left<\Delta\left<[O_3]\right>\right>} derived for Svanvik in 2018 is slightly larger than for Esrange and Pallas. This could mean either that 2018 had been more exceptional at Svanvik or the Svanvik climatology may need a correction for increased tropospheric background ozone concentration. Looking at daily \chem{\Delta_{present-past}\left<[O_3]\right>}, wherein present is defined as the average of 2010--2018 and past as the respective beginning of data taking until 1997, does not reveal a clear trend: We find for Esrange $(1.5\pm 3.1)\,\unit{ppb}$ and for Pallas $(-2.5\pm 3.7)\,\unit{ppb}$.

To further investigate the ozone residuals, we conduct a Student's~t-test ($t = \frac{\chem{\Delta[O_3]}}{\sigma}$), assuming the same sample uncertainty $\sigma$ in both, climatology and the year in question. Results for Esrange and Pallas are shown in Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test}. This t-test clearly shows, that at both sites for at least $50\,\unit{\%}$ of days the residuals are significant on the $1\sigma$ level. In Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}, the test statistics for Svanvik are displayed. From Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}a), we can deduce, that the residuals of the Svanvik climatology from the northern Fennoscandic climatology are significant on the $2\sigma$ level for at least half of the days, while the residuals are less certain in 2018 and 2019. In Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}b) the same is shown for the years 2018/19 with respect to the climatology from Svanvik. We find that in less than $50\,\unit{\%}$ of days the ozone concentration residuals is significant on the $1\sigma$ level.

@@

\appendix
\section{}    %% Appendix A

\subsection{}     %% Appendix A1, A2, etc.

@@

\appendixfigures  %% needs to be added in front of appendix figures

\begin{figure}[t]
  \includegraphics[width=8.3cm]{global_rad_clim}
  \caption{Observed global irradiance at Svanvik. The cyan lines indicate the $1\,\sigma$ level deducted from the climatology. (a) Climatology (hourly maxima, mean, and minima); deviation from climatology (b) 2018; (c) 2019.}
  \label{fig:global_rad_clim}
\end{figure}

\begin{figure}[t]
  \includegraphics[width=8.3cm]{test_fennoscandia}
  \caption{Maximum daily sun shine duration based on geometrical calculations at different sites in Fennoscandia. Midnight sun conditions at Jergul/Karasjok and Svanvik prevail from the end of May until the end of July, while at Esrange and Pallas they only prevail from the beginning of June until mid of July.}
  \label{fig:sunlight_fennoscandia}
\end{figure}

\begin{figure}[t]
  \includegraphics[width=8.3cm]{ozone_obs_2018_07}
  \caption{Estimate of missing observation \chem{[O_3]} data at Svanhovd in July 2018 based on derived hourly climatologies of northern Fennoscandia, Svanvik, and Copernicus regional model reanalysis ensemble mean. (a) Time series; (b) Student's t-test.}
  \label{fig:ozone_obs_2018_rrea}
\end{figure}

\begin{figure}[t]
  \includegraphics[width=8.3cm]{greening_season_change_Svanvik}
  \caption{Estimated shift and prolongation of greening season at Svanhovd over the past two decades based on the method and data described in \citet{GMD:Falk2019}.}
  \label{fig:greening_season_change_Svanvik}
\end{figure}

\appendixtables   %% needs to be added in front of appendix tables

>>>>

The \chem{\left<\Delta\left<[O_3]\right>\right>} derived for Svanvik in 2018 is slightly larger than for Esrange and Pallas. This could mean either that 2018 had been more exceptional at Svanvik or the Svanvik climatology may need a correction for increased tropospheric background ozone concentration. Based on the meteorological conditions, we may assume 2019 as an average year which would lead to a deduction of a constant bias of about $1.2\,\unit{ppb}$ in the Svanvik climatology. Daily \chem{\Delta_{present-past}\left<[O_3]\right>} from Esrange and Pallas, wherein present is defined as the average of 2010--2018 and past as the respective beginning of data taking until 1997, do not reveal a clear trend: We find for Esrange $(1.5\pm 3.1)\,\unit{ppb}$ and for Pallas $(-2.5\pm 3.7)\,\unit{ppb}$. Former is compatible with our assumed bias for the Svanvik climatology.

To further investigate the ozone residuals, we conduct a Student's~t-test ($t = \frac{\chem{\Delta[O_3]}}{\sigma}$), assuming the same sample uncertainty $\sigma$ in both, climatology and the year in question. Results for Esrange and Pallas are shown in Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test}. This t-test clearly shows, that at both sites for at least $50\,\unit{\%}$ of days the residuals are significant on the $1\sigma$ level. In Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}, the test statistics for Svanvik are displayed. From Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}a), we can deduce, that the residuals of the Svanvik climatology from the northern Fennoscandic climatology are significant on the $2\sigma$ level for at least half of the days, while the residuals are less certain in 2018 and 2019. In Fig.~\ref{fig:ozone_climatology_fenoscandic_obs_test-Svanvik}b) the same is shown for the years 2018/19 with respect to the climatology from Svanvik. We find that in less than $50\,\unit{\%}$ of days the ozone concentration residuals are significant on the $1\sigma$ level.

\begin{figure}[t]
  \includegraphics[width=8.3cm]{ozone_signific}
  \caption{Ozone concentration anomalies at Svanvik. Percentage of days above/below $\pm 1\,\sigma$ from climatological mean displayed for each month. Negative deviations from climatology are shown as negative percentage. The annual positive/negative deviations are indicated in the respective corners (right upper/right lower). The dashed lines indicate the expected percentage of values falling above/below $\pm 1\,\sigma$ if a normal distribution is assumed ($15.9\,\unit{\%}$). Year displayed in panel (a) 2018; (b) 2019. In April/May 2018, ozone was significantly higher than on average, while 2019 was generally an average year. In both years ozone concentrations were significantly below average in November.}
  \label{fig:ozone_signific}
\end{figure}

@@

\appendix
\section{Derived climatologies}    %% Appendix A

\subsection{Temperature, precipitation, and global irradiance}     %% Appendix A1, A2, etc.

\appendixfigures  %% needs to be added in front of appendix figures
\begin{figure}[t]
  \includegraphics[width=8.3cm]{global_rad_clim}
  \caption{Observed global irradiance at Svanvik. The cyan lines indicate the $1\,\sigma$ level deducted from the climatology. (a) Climatology (hourly maxima, mean, and minima); deviation from climatology (b) 2018; (c) 2019.}
  \label{fig:global_rad_clim}
\end{figure}
\clearpage
\begin{figure}[t]
  \includegraphics[width=8.3cm]{test_fennoscandia}
  \caption{Maximum daily sun shine duration based on geometrical calculations at different sites in Fennoscandia. Midnight sun conditions at Jergul/Karasjok and Svanvik prevail from the end of May until the end of July, while at Esrange and Pallas they only prevail from the beginning of June until mid of July.}
  \label{fig:sunlight_fennoscandia}
\end{figure}
\clearpage
\begin{figure}[t]
  \includegraphics[width=8.3cm]{greening_season_change_Svanvik}
  \caption{Estimated shift and prolongation of greening season at Svanhovd over the past two decades based on the method and data described in \citet{GMD:Falk2019}.}
  \label{fig:greening_season_change_Svanvik}
\end{figure}
\clearpage
\subsection{Ozone}

@@

%\appendixfigures  %% needs to be added in front of appendix figures

%\begin{figure}[t]
%  \includegraphics[width=8.3cm]{ozone_obs_2018_07}
%  \caption{Estimate of missing observation \chem{[O_3]} data at Svanhovd in July 2018 based on derived hourly climatologies of northern Fennoscandia, Svanvik, and Copernicus regional model reanalysis ensemble mean. (a) Time series; (b) Student's t-test.}
%  \label{fig:ozone_obs_2018_rrea}
%\end{figure}

%\noappendix

%\appendixtables   %% needs to be added in front of appendix tables
